Compararison of analytical and artificial neural network models in predicting hot deformation behavior of AISI P20 and interpolation-extrapolation capacity validation

### 2022-11-10 Article.tex
Add Appendix to the paper

Once the learning phase is over, the trained model can be used to predict the behavior of the AISI P20 alloy as a function of the input data in the same way as is done with the analytical models.
One can either use the model directly by providing it with new input data, or retrieve the $180$ parameters identified during the training and inject them into a mathematical model based on Equations (\ref{eq:ANN1}) and (\ref{eq:ANN2}) which can be implemented in any language (\eg in FORTRAN for use on the Abaqus Explicit FEM code) as proposed in Pantalé \eal \cite{Pantale-2021}.
For reasons of compactness, the parameters of the ANN model are not reported in this paper.

@@

%---------------------------------------------------------------------
\section{Conclusion \label{sec:Conclusion}}
Experiments were conducted on modified AISI P20 carbon alloy to study the applicability and predictive accuracy of five analytical models and one neural network model over a range of temperatures ($1050\celsius$ - $1250\celsius$, strains ($0.025$ - $0.7$), strain rates ($0.001\ \ps$ - $5\ \ps$) and the following conclusions were drawn.
\begin{enumerate}
\item The flow stress increases with decreasing temperature and increasing strain rate due to the competitive occurrence of dynamic softening and strain hardening mechanisms.
The degree of DRX was discussed through the difference between the maximum and permanent stress, as well as the evolution of the microstructure which shows a partially complete DRX.
At high strain rates it is difficult to visualise the DRX phenomenon on the flow curves due to the strain rate sensitivity of this phenomenon.
A further study will focus on deep analysis of the microstructure of this steel alloy and its impact on mechanical properties.
\item Five analytical models have been identified on this alloy and a model based on neural networks.
Among the analytical models, the JC, HS and MZA models proved inappropriate for analysing the behavior of this material whereas the PTM and AR models showed their capacity since they have a relatively acceptable error ($ 3.2 \% $ and $ 3.8 \%$).
As for the ANN model, it is largely reliable that the analytical models to predict the competitiveness of AISI P20.
Therefore it may be advisable to use the ANN model when conducting a study involving this alloy under these experimental conditions.
\item To test the performance of each model, a study is carried out to evaluate the interpolation and extrapolation capacity of the developed models.
For the interpolation framework, all models have a good correlation with the experiment except the HS model but the ANN model shows a better correlation.
For extrapolation of the data, the analytical models do not show good predictive capacity compared to the ANN model which still shows good extrapolation capacity.
This further confirms what was demonstrated during his identification.
\item As perspectives of this work, we will develop also an ANN model to characterize the microstructure in terms of DRX and phase formation such as fearrite, bainite and martensite as an example.
This ANN model can therefore be compared to the most commonly used classical models such as the JMAK model.
Therfore, this ANN model  will be implemented in FEM softwares such as Abaqus, Forge in order to validate it by doing a simulation.
\end{enumerate}
% \bibitem{}
\bibliographystyle{elsarticle-num}
\bibliography{Bibliography}
%\end{thebibliography}
\end{document}
\endinput

>>>>

Once the learning phase is over, the trained model can be used to predict the behavior of the AISI P20 alloy as a function of the input data in the same way as is done with the analytical models.
One can either use the model directly by providing it with new input data, or retrieve the $180$ parameters identified during the training and inject them into a mathematical model based on Equations (\ref{eq:ANN1}) and (\ref{eq:ANN2}) which can be implemented in any language (\eg in FORTRAN for use on the Abaqus Explicit FEM code) as proposed in Pantalé \eal \cite{Pantale-2021}.
For reasons of compactness, the parameters of the ANN model and the complete procedure to compute the flow stress $\sigma^y$ from the input data are provided in the Appendix \ref{sec:Appendix} at the end of this paper.

@@

%---------------------------------------------------------------------
\section{Conclusion \label{sec:Conclusion}}
Experiments were conducted on modified AISI P20 carbon alloy to study the applicability and predictive accuracy of five analytical models and one neural network model over a range of temperatures ($1050\celsius$ - $1250\celsius$, strains ($0.025$ - $0.7$), strain rates ($0.001\ \ps$ - $5\ \ps$) and the following conclusions were drawn.
\begin{enumerate}
\item The flow stress increases with decreasing temperature and increasing strain rate due to the competitive occurrence of dynamic softening and strain hardening mechanisms.
The degree of DRX was discussed through the difference between the maximum and permanent stress, as well as the evolution of the microstructure which shows a partially complete DRX.
At high strain rates it is difficult to visualise the DRX phenomenon on the flow curves due to the strain rate sensitivity of this phenomenon.
A further study will focus on deep analysis of the microstructure of this steel alloy and its impact on mechanical properties.
\item Five analytical models have been identified on this alloy and a model based on neural networks.
Among the analytical models, the JC, HS and MZA models proved inappropriate for analysing the behavior of this material whereas the PTM and AR models showed their capacity since they have a relatively acceptable error ($ 3.2 \% $ and $ 3.8 \%$).
As for the ANN model, it is largely reliable that the analytical models to predict the competitiveness of AISI P20.
Therefore it may be advisable to use the ANN model when conducting a study involving this alloy under these experimental conditions.
\item To test the performance of each model, a study is carried out to evaluate the interpolation and extrapolation capacity of the developed models.
For the interpolation framework, all models have a good correlation with the experiment except the HS model but the ANN model shows a better correlation.
For extrapolation of the data, the analytical models do not show good predictive capacity compared to the ANN model which still shows good extrapolation capacity.
This further confirms what was demonstrated during his identification.
\item As perspectives of this work, we will develop also an ANN model to characterize the microstructure in terms of DRX and phase formation such as fearrite, bainite and martensite as an example.
This ANN model can therefore be compared to the most commonly used classical models such as the JMAK model.
Therfore, this ANN model  will be implemented in FEM softwares such as Abaqus, Forge in order to validate it by doing a simulation.
\end{enumerate}

\section{Appendix\label{sec:Appendix}}
In order to complete this paper, we report here after the computing process and the $180$ coefficients of the Artificial Neural Network ANN-3-15-7-1-sigmoid model used in Section \ref{sec:ANNmodel}. In order to use this model we describe here after the procedure to compute the flow stress $\sigma^y$ from the input variables $\varepsilon^p$, $\mdot\varepsilon$ and $T$. This process can be decomposed in $3$ phases:
\begin{itemize}
\item We first have to normalize the input values of the ANN $x_i$ within the range $[0,1]$ to avoid ill-conditioned system as presented by many other authors in the literature \cite{Lin-2008-ANN, Lu-2011-ANN}.
Therefore, the three components of the input are obtained from the plastic strain $\varepsilon^p$, the plastic strain rate $\mdot{\varepsilon}^p$ and the temperature $T$ using the following expressions:
\begin{equation}
\begin{cases}
x_1 = \frac{\varepsilon^p - [\varepsilon^p]_{min}}{[\varepsilon^p]_{max} - [\varepsilon^p]_{min}}\\
x_2 = \frac{\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})-[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{min}}{[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{max}-[\ln(\mdot{\varepsilon}/\mdot{\varepsilon_0})]_{min}}\label{eq:CR1}\\
x_3 = \frac{T-[T]_{min}}{[T]_{max}-[T]_{min}}
\end{cases}
\end{equation}
where $[~]_{min}$ and $[~]_{max}$  are the boundaries of the range of the corresponding field: $\varepsilon^p\!\in\!\left[0.0,0.7\right]$, $\mdot{\varepsilon}\!\in\!\left[0.001,5.0\right]$, $T\!\in\!\left[1050,1250\right]$ and $\sigma\!\in\!\left[1.311,153.739\right]$. The reference strain rate is $\mdot{\varepsilon_0} = 0.001$.
\item Then we compute the output $s$ of the ANN from the input vector $\overrightarrow{x}$ using the three following equations.
\begin{equation}
\overrightarrow{y}_1 = \left[1 + \exp{\left(- \w_1 \dotp \overrightarrow{x}- \overrightarrow{b}_1\right)}\right]^{-1}
\end{equation}
\begin{equation}
\overrightarrow{y}_2 = \left[1 + \exp{\left(- \w_2 \dotp \overrightarrow{y}_1- \overrightarrow{b}_2\right)}\right]^{-1}
\end{equation}
\begin{equation}
s = \overrightarrow{w}^T \dotp \overrightarrow{y}_2 + b
\end{equation}
\item Finally, the flow stress $\sigma^y$ can be obtained from the output $s$ of the ANN using the following equation:
\begin{equation}
\sigma^y =  \left([\sigma]_{max}-[\sigma]_{min}\right)s + [\sigma]_{min} \label{eq:CR2}
\end{equation}
\end{itemize}

Conforming to the computing process proposed by Equations (\ref{eq:CR1}-\ref{eq:CR2}), we report hereafter the $180$ coefficients of the ANN-3-15-7-1-sigmoid model used in Section \ref{sec:ANNmodel}.

\begin{equation*}
\w_1^T = \left[
\begin{array}{rrr}
2.2206 & -3.7555 & -6.7246\\ 
-4.8598 & 5.7431 & -5.8538\\ 
2.3099 & 3.3325 & -5.1795\\ 
2.0475 & 0.8006 & -1.4259\\ 
8.8358 & -6.0362 & 0.8226\\ 
-1.2613 & -0.9274 & -2.3725\\ 
-0.3561 & 6.5032 & -10.7573\\ 
-11.7226 & -2.0455 & 1.2248\\ 
3.1066 & 26.5580 & 18.6540\\ 
-0.5150 & -5.6922 & 1.0104\\ 
-6.4755 & 8.4888 & -2.4459\\ 
-1.8791 & -0.5380 & 2.2295\\ 
-6.0206 & 1.2776 & 0.2169\\ 
0.2619 & -4.7974 & -1.1282\\ 
-27.1456 & -0.5327 & -0.5303\\ 
\end{array}\right]
\end{equation*}
\begin{equation*}
\overrightarrow{b}_1 = \left[
\begin{array}{r}
5.9164\\ 
-1.9074\\ 
-2.6837\\ 
-1.0954\\ 
-0.9509\\ 
3.1682\\ 
-4.3964\\ 
0.6343\\ 
-5.0676\\ 
2.0228\\ 
1.1267\\ 
-0.9671\\ 
-0.6263\\ 
2.3110\\ 
-0.3037\\ 
\end{array}\right]
\end{equation*}
\begin{equation*}
\w_2 = \left[
\begin{array}{rrrrrrr}
-0.5783 & 1.2724 & 0.5747 & 0.6449 & -4.2203 & -0.2380 & 0.2591\\ 
-0.4852 & 5.2807 & 0.8888 & -8.2324 & -2.0075 & -0.6474 & -1.0787\\ 
4.4499 & -0.0137 & 0.1657 & 0.3198 & 4.9765 & -1.2503 & 0.8219\\ 
1.7571 & 0.7730 & 0.0208 & -1.3316 & -0.8945 & -0.7284 & -0.1831\\ 
-1.0866 & 0.1330 & -0.8615 & -0.1283 & 0.2218 & -0.1772 & -2.7458\\ 
-0.3925 & 1.3994 & 0.0630 & -1.8397 & -1.1047 & -1.9839 & 0.5767\\ 
-0.2121 & 0.9977 & 1.2028 & -9.6525 & 0.5520 & 0.1062 & -0.0409\\ 
-1.1518 & -1.6402 & -4.1501 & -1.0759 & 0.4749 & -2.8350 & 0.9225\\ 
-1.1453 & -0.2173 & -0.1382 & 0.8264 & -0.5125 & 0.1882 & -0.8654\\ 
-4.3301 & -0.3711 & -7.4305 & 3.5926 & -9.6217 & -1.2375 & 1.6171\\ 
2.3907 & -1.0085 & -0.8828 & -1.1891 & 0.9947 & 1.1178 & -1.0953\\ 
-1.5955 & 1.6313 & 0.4916 & 0.1906 & -1.9216 & -1.4140 & 1.3827\\ 
-1.6985 & 1.4277 & -3.4462 & -8.3777 & -1.2132 & -1.2158 & 2.8512\\ 
-1.9954 & -2.0159 & -8.3455 & -0.5205 & 0.2942 & -1.3337 & 0.2026\\ 
4.3040 & -0.7164 & -1.0859 & 3.4294 & -23.8003 & 12.5859 & 7.3721\\ 
\end{array}\right]
\end{equation*}
\begin{equation*}
\overrightarrow{b}_2 = \left[
\begin{array}{r}
0.7534\\ 
0.9473\\ 
0.6055\\ 
-0.7793\\ 
-1.0305\\ 
-1.5779\\ 
-0.1471\\ 
\end{array}\right]
\end{equation*}
\begin{equation*}
\overrightarrow{w}_3 = \left[
\begin{array}{r}
0.1920\\ 
0.3406\\ 
0.3839\\ 
-0.2880\\ 
1.2047\\ 
-1.4126\\ 
-0.2215\\ 
\end{array}\right]
\end{equation*}
\begin{equation*}
b_3 = 0.1178 
\end{equation*}

% \bibitem{}
\bibliographystyle{elsarticle-num}
\bibliography{Bibliography}
%\end{thebibliography}
\end{document}
\endinput
